### 2016-03-26 ch05-initvals/initial-values.tex
LPM illustration improved

\section{Introduction}
\label{se:introd}
The task of interpreting measurement data from dynamic systems often involves the estimation of a transfer function (TF). 
Typical measurement techniques and identification strategies of  \glspl{TF} or \glspl{FRF} are presented in~\citep{Schoukens1998,Schoukens2006LPM,Guillaume1996,Broersen1995,Pintelon2010LPM1,Antoni2007FRF}.
Application of these methods to real devices and systems is well-represented in the literature, e.g. \citep{Lim2010,Robinson1990,Behjat2010}.

@@

\paragraph{Outline}
The rest of the chapter is structured as follows. 
\secref{se:ProbForm} covers the problem formulation. 
\secref{se:MethodEg} presents the methodology for obtaining the initial estimates, and their influence on the success rate of the \gls{MLE}. 
This is followed by \secref{se:Demo} and \secref{se:ExpMeas}, which are demonstrations of the improved initial estimates on simulation and experimental data, respectively.
In \secref{se:Generality} the generality of the results is investigated briefly.
The computational effort and recommendations for practical implementations are given in \secref{sec:initvals:computation}.
Ultimately, concluding remarks are presented in \secref{se:Conclusion}.

\section{Problem formulation}\label{se:ProbForm}

@@

\subsection{System Setup Framework and Model}
\TODO{ hele subsectie verplaatsen na het param. sys. model?}
\figref{fig:oesetup} depicts a schematic of the output error framework for  a generalized (single- or multiple-order) resonating, dynamic \gls{LTI} discrete-time \gls{SISO} system, subjected to a known white random noise input. 
The full model of the system is
\begin{equation}\label{lpmtd1}
y(t)=\true{G}(q^{-1})u_0(t)+H_0(q^{-1})e(t)
\end{equation}
where $\true{G}(q^{-1})$ represents the dynamics of the system to be estimated, $u_0(t)$ is the input signal, $v(t)= H_0(q^{-1})e(t)$ is the noise source at the output, $H_0(q^{-1})$ is the noise dynamics, 
$e(t)$ is white Gaussian noise, and $q^{-m}$ is the backwards shift operator ($q^{-m}x(t)$ = $x(t-mT_{\mathrm{s}})$  with $m$ a positive integer and $T_{\mathrm{s}}$ the sampling time).
We only treat the discrete-time case (i.e. $t = n \cdot T_{\mathrm{s}}$ for integer values of $n$) theoretically in this chapter.
However, the generalization to continuous time is straightforward~\citep[Chapter 6]{Pintelon2012} and has been demonstrated in Section~\ref{se:ExpMeas}.

@@

In this chapter we assume that the output signal is disturbed by random noise, resulting in noisy error-prone data.
For the simulations, white noise is used ($\true{H} = 1$ and $e(t) = v(t)$).
It is also possible to apply the estimation procedure in this chapter to a system that is disturbed by colored noise, as will be demonstrated in \secref{se:ExpMeas}.

@@

With reference to equation \eqref{lpmtd1}, the relation between the noiseless input and the output signals ($v(t)= 0$) is assumed to be of the form
\begin{equation}
    A(q^{-1}) y_0(t) = B(q^{-1}) u_0(t) \quad \forall t \in n \Ts{}  \text{ with }  n \in \IntegerNumbers
\label{ABpolys}
\end{equation}
where $A$ and $B$ are polynomials in $q^{-1}$. 
Thus, it follows from equation \eqref{lpmtd1} that
\begin{equation}
    \true{G}(q^{-1})= \frac{B(q^{-1})}{A(q^{-1})}
    \text{.}
\end{equation}

@@

\subsection{Parametric Identification Algorithm}\label{se:paramIdentAlgo}

@@

\section{Methodology}\label{se:MethodEg}
In this section, the considered methods for obtaining the initial estimates of $\theta$ are briefly explained. Then their influence on the success rate of the maximum likelihood estimator is described.
%}

@@

%The latter 2 ....schematic is configured to include connections of the initial values and the input to the LPM block and/or to the MLE block(s) for estimation purposes.
%\end{itemize}
%\section{Example}\label{se:Example}

@@

The $\mathcal{L}_2$ norm used in equation \eqref{eq:successrateTol} is defined~\citep{Skogestad2005} as
\begin{align}
\norm[2]{G(z)} &\isdef \sqrt{\frac{1}{2\pi}\int_0^{2\pi} |G(e^{-j\omega})|^2\ \mathrm{d}\omega}& \text{ for discrete time,}\\
\norm[2]{G(s)} &\isdef \sqrt{\frac{1}{2\pi}\int_{-\infty}^{+\infty} |G(j\omega)|^2\ \mathrm{d}\omega}& \text{ for continuous time}
\end{align}
 which, in practice, is computed by using the adaptive global quadrature algorithm provided by the MATLAB function \code{integral}.
The success rate in equation \eqref{eq:successrateTol} is estimated via Monte Carlo simulations in \secref{se:CompuSR}.

@@

\section{Demonstration}\label{se:Demo}

@@

%\subsection{Computing the success rates}
\subsection{Computation of  the Success Rates}\label{se:CompuSR}

@@

\section{Experimental Results}
\label{se:ExpMeas}
Measurements were performed on an electrical filter to assess the performance of the initial value generating methods in a physical setting.

@@

\section{Remark on the generality}
\label{se:Generality}
In the previous sections, the usability of the smoothers to produce initial values has been studied on very specific examples.
In this section, an attempt will be made to illustrate the usefulness of these smoothers for the generation of starting values for other systems.
However, due to the intricate relationship between the attraction regions in the cost function, the location of the actual system poles, the \gls{SNR} level and the choice of a particular excitation signal, we think it is intractable to construct rigid bounds on where the smoothers produce effective starting values.

@@

To measure the timing of the different methods, essentially the same settings were used as in \secref{se:Demo}: the systems from equations \eqref{eq:systemundertest} and \eqref{eq:systundertest-20dB} were simulated such that $N_{3\unit{dB}} \in \Set{36,48}$ and $\SNR = 0 \unit{dB}$.
For each of the branches in \figref{fig:flowgraph}, three different timings are recorded (when applicable):
\begin{itemize}
  \item $t_{\mathrm{smooth}}$ is the time it takes to perform the smoothing (all blocks before \gls{GTLS}/\gls{BTLS}),
  \item $t_{\mathrm{init}}$ is the time it takes to fit $\model[init]{\bullet}$, and
  \item $t_{\mathrm{final}}$ is the time it takes to fit $\model{\bullet}$.
\end{itemize}
The median timings as returned by \mcode{timeit} are reported in \tabref{tbl:init:timing}.

@@

\section{Conclusion}\label{se:Conclusion}
The simulations have demonstrated that use of a smoothed non-parametric estimate of the \gls{FRF} can improve the success rate of minimizing the \gls{ML} cost function for the estimation of the transfer functions of \gls{LTI} systems, subjected to noisy signals. Specifically, two smoothing methods were used and compared: 1) the truncated \gls{LPM} method, and 2) the \gls{RFIR} method. The simulation results clearly show that the \gls{RFIR} method is superior to the truncated \gls{LPM} method and the existing \gls{BTLS} and \gls{GTLS} methods.

>>>>

\section{Introduction}
\label{sec:initial-values:introd}
The task of interpreting measurement data from dynamic systems often involves the estimation of a transfer function (TF). 
Typical measurement techniques and identification strategies of  \glspl{TF} or \glspl{FRF} are presented in~\citep{Schoukens1998,Schoukens2006LPM,Guillaume1996,Broersen1995,Pintelon2010LPM1,Antoni2007FRF}.
Application of these methods to real devices and systems is well-represented in the literature, e.g. \citep{Lim2010,Robinson1990,Behjat2010}.

@@

\paragraph{Outline}
The rest of the chapter is structured as follows. 
\secref{sec:initial-values:ProbForm} covers the problem formulation. 
\secref{sec:initial-values:MethodEg} presents the methodology for obtaining the initial estimates, and their influence on the success rate of the \gls{MLE}. 
This is followed by \secref{sec:initial-values:Demo} and \secref{sec:initial-values:ExpMeas}, which are demonstrations of the improved initial estimates on simulation and experimental data, respectively.
In \secref{sec:initial-values:Generality} the generality of the results is investigated briefly.
The computational effort and recommendations for practical implementations are given in \secref{sec:initvals:computation}.
Ultimately, concluding remarks are presented in \secref{sec:initial-values:Conclusion}.

\section{Problem formulation}\label{sec:initial-values:ProbForm}

@@

\subsection{System Setup Framework and Model}
\TODO{ hele subsectie verplaatsen na het param. sys. model?}
\figref{fig:oesetup} depicts a schematic of the output error framework for  a generalized (single- or multiple-order) resonating, dynamic \gls{LTI} discrete-time \gls{SISO} system, subjected to a known white random noise input. 
The full model of the system is
\begin{equation}
y(t)=\true{G}(q^{-1})u_0(t)+H_0(q^{-1})e(t)
\label{eq:initial-values:OE:TD}
\end{equation}
where $\true{G}(q^{-1})$ represents the dynamics of the system to be estimated, $u_0(t)$ is the input signal, $v(t)= H_0(q^{-1})e(t)$ is the noise source at the output, $H_0(q^{-1})$ is the noise dynamics, 
$e(t)$ is white Gaussian noise, and $q^{-m}$ is the backwards shift operator ($q^{-m}x(t)$ = $x(t-mT_{\mathrm{s}})$  with $m$ a positive integer and $T_{\mathrm{s}}$ the sampling time).
We only treat the discrete-time case (i.e. $t = n \cdot T_{\mathrm{s}}$ for integer values of $n$) theoretically in this chapter.
However, the generalization to continuous time is straightforward~\citep[Chapter 6]{Pintelon2012} and has been demonstrated in Section~\ref{sec:initial-values:ExpMeas}.

@@

In this chapter we assume that the output signal is disturbed by random noise, resulting in noisy error-prone data.
For the simulations, white noise is used ($\true{H} = 1$ and $e(t) = v(t)$).
It is also possible to apply the estimation procedure in this chapter to a system that is disturbed by colored noise, as will be demonstrated in \secref{sec:initial-values:ExpMeas}.

@@

With reference to equation \eqref{eq:initial-values:OE:TD}, the relation between the noiseless input and the output signals ($v(t)= 0$) is assumed to be of the form
\begin{equation}
    A(q^{-1}) y_0(t) = B(q^{-1}) u_0(t) \quad \forall t \in n \Ts{}  \text{ with }  n \in \IntegerNumbers
\label{ABpolys}
\end{equation}
where $A$ and $B$ are polynomials in $q^{-1}$. 
Thus, it follows from equation \eqref{eq:initial-values:OE:TD} that
\begin{equation}
    \true{G}(q^{-1})= \frac{B(q^{-1})}{A(q^{-1})}
    \text{.}
\end{equation}

@@

\subsection{Parametric Identification Algorithm}\label{sec:initial-values:paramIdentAlgo}

@@

\section{Methodology}\label{sec:initial-values:MethodEg}
In this section, the considered methods for obtaining the initial estimates of $\theta$ are briefly explained. Then their influence on the success rate of the maximum likelihood estimator is described.
%}

@@

%The latter 2 ....schematic is configured to include connections of the initial values and the input to the LPM block and/or to the MLE block(s) for estimation purposes.
%\end{itemize}
%\section{Example}\label{sec:initial-values:Example}

@@

The $\mathcal{L}_2$ norm used in equation \eqref{eq:successrateTol} is defined~\citep{Skogestad2005} as
\begin{align}
\norm[2]{G(z)} &\isdef \sqrt{\frac{1}{2\pi}\int_0^{2\pi} |G(e^{-j\omega})|^2\ \mathrm{d}\omega}& \text{ for discrete time,}\\
\norm[2]{G(s)} &\isdef \sqrt{\frac{1}{2\pi}\int_{-\infty}^{+\infty} |G(j\omega)|^2\ \mathrm{d}\omega}& \text{ for continuous time}
\end{align}
 which, in practice, is computed by using the adaptive global quadrature algorithm provided by the MATLAB function \code{integral}.
The success rate in equation \eqref{eq:successrateTol} is estimated via Monte Carlo simulations in \secref{sec:initial-values:CompuSR}.

@@

\section{Demonstration}\label{sec:initial-values:Demo}

@@

%\subsection{Computing the success rates}
\subsection{Computation of  the Success Rates}\label{sec:initial-values:CompuSR}

@@

\section{Experimental Results}
\label{sec:initial-values:ExpMeas}
Measurements were performed on an electrical filter to assess the performance of the initial value generating methods in a physical setting.

@@

\section{Remark on the generality}
\label{sec:initial-values:Generality}
In the previous sections, the usability of the smoothers to produce initial values has been studied on very specific examples.
In this section, an attempt will be made to illustrate the usefulness of these smoothers for the generation of starting values for other systems.
However, due to the intricate relationship between the attraction regions in the cost function, the location of the actual system poles, the \gls{SNR} level and the choice of a particular excitation signal, we think it is intractable to construct rigid bounds on where the smoothers produce effective starting values.

@@

To measure the timing of the different methods, essentially the same settings were used as in \secref{sec:initial-values:Demo}: the systems from equations \eqref{eq:systemundertest} and \eqref{eq:systundertest-20dB} were simulated such that $N_{3\unit{dB}} \in \Set{36,48}$ and $\SNR = 0 \unit{dB}$.
For each of the branches in \figref{fig:flowgraph}, three different timings are recorded (when applicable):
\begin{itemize}
  \item $t_{\mathrm{smooth}}$ is the time it takes to perform the smoothing (all blocks before \gls{GTLS}/\gls{BTLS}),
  \item $t_{\mathrm{init}}$ is the time it takes to fit $\model[init]{\bullet}$, and
  \item $t_{\mathrm{final}}$ is the time it takes to fit $\model{\bullet}$.
\end{itemize}
The median timings as returned by \mcode{timeit} are reported in \tabref{tbl:init:timing}.

@@

\section{Conclusion}\label{sec:initial-values:Conclusion}
The simulations have demonstrated that use of a smoothed non-parametric estimate of the \gls{FRF} can improve the success rate of minimizing the \gls{ML} cost function for the estimation of the transfer functions of \gls{LTI} systems, subjected to noisy signals. Specifically, two smoothing methods were used and compared: 1) the truncated \gls{LPM} method, and 2) the \gls{RFIR} method. The simulation results clearly show that the \gls{RFIR} method is superior to the truncated \gls{LPM} method and the existing \gls{BTLS} and \gls{GTLS} methods.
